---
title: "Project 2 - Alon Goodman & Ran Hassid"
output:
  html_document:
    df_print: paged
---

# Q1 #

```{r}
#install.packages("igraph")
library(igraph)
```

### The System ###

```{r,echo=FALSE}
graph.Q1 <- make_graph(c(
  "A", "B", "A", "C","A", "D","A", "E",
  "B","C","B","F",
  "C","F","C","G","C","D","C","H",
  "D","E","D","H","D","I","D","G",
  "E","I",
  "F","G","F","J",
  "G","H","G","J",
  "H","I","H","J",
  "I","J"), directed = TRUE)

plot(graph.Q1)
```

## a ##

The distribution of the time of until an edge will not working is exp(theta).

So, the probability that an edge will not work after 10 days:

P(exp(theta)<10) = 1 - e^(-10 * theta)

```{r}
theta <- 0.02

h_x.result_vec.a <- numeric(15000)

for (i in 1:15000) {
  graph.zero_day.i <- make_graph(c("A", "B", "A", "C","A", "D","A", "E",
                           "B","C","B","F",
                           "C","F","C","G","C","D","C","H",
                           "D","E","D","H","D","I","D","G",
                           "E","I",
                           "F","G","F","J",
                           "G","H","G","J",
                           "H","I","H","J",
                           "I","J"), directed = TRUE)
  after_10_days.i <- rbinom(n = 22,size = 1,prob = 1-exp(-10*theta))
  graph_after_10_days.i <- delete.edges(graph = graph.zero_day.i,edges = which(after_10_days.i==1))
  paths_A_J_after_10_days.i <- all_simple_paths(graph = graph_after_10_days.i,from = "A","J")
  if(identical(paths_A_J_after_10_days.i,list())){
    h_x.result_vec.a[i] <- 1
  }
  
}

```


```{r,echo=FALSE}
cat(sprintf("Expected value estimator = %s", round(mean(h_x.result_vec.a),4)),
    sprintf("Varinace estimator = %s", round(var(h_x.result_vec.a),4)),
    sep = "\n")
```


```{r}

est.int <- cumsum(h_x.result_vec.a)/(1:15000)
est.err <- sqrt( cumsum( (h_x.result_vec.a-est.int)^2 )) / (1:15000)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2, ylim = mean(h_x.result_vec.a)+100*est.err[15000]*c(-1,1))
lines(est.int + 2*est.err, col = "gold", lwd = 2)
lines(est.int - 2*est.err, col = "gold", lwd = 2)
```
 
## b ##

```{r}
theta <- 0.005

h_x.result_vec.b <- numeric(50000)

for (i in 1:50000) {
  graph.zero_day.i <- make_graph(c("A", "B", "A", "C","A", "D","A", "E",
                           "B","C","B","F",
                           "C","F","C","G","C","D","C","H",
                           "D","E","D","H","D","I","D","G",
                           "E","I",
                           "F","G","F","J",
                           "G","H","G","J",
                           "H","I","H","J",
                           "I","J"), directed = TRUE)
  after_10_days.i <- rbinom(n = 22,size = 1,prob = 1-exp(-10*theta))
  graph_after_10_days.i <- delete.edges(graph = graph.zero_day.i,edges = which(after_10_days.i==1))
  paths_A_J_after_10_days.i <- all_simple_paths(graph = graph_after_10_days.i,from = "A","J")
  if(identical(paths_A_J_after_10_days.i,list())){
    h_x.result_vec.b[i] <- 1
  }
  
}

```

```{r,echo=FALSE}
cat(sprintf("Expected value estimator = %s", mean(h_x.result_vec.b)),
    sprintf("Varinace estimator = %s", var(h_x.result_vec.b)),
    sprintf("Out of 50,000 test, the no. of failures = %s", sum(h_x.result_vec.b)),
    sep = "\n")
```

```{r}

est.int <- cumsum(h_x.result_vec.b)/(1:50000)
est.err <- sqrt( cumsum( (h_x.result_vec.b-est.int)^2 )) / (1:50000)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2, ylim = mean(h_x.result_vec.b)+100*est.err[50000]*c(-1,1))
lines(est.int + 2*est.err, col = "gold", lwd = 2)
lines(est.int - 2*est.err, col = "gold", lwd = 2)
```

## c ## 

```{r}
theta.f <- 0.02
theta.g <- 0.05

h_x.result_vec.c <- numeric(15000)

for (i in 1:15000) {
  graph.zero_day.i <- make_graph(c("A", "B", "A", "C","A", "D","A", "E",
                           "B","C","B","F",
                           "C","F","C","G","C","D","C","H",
                           "D","E","D","H","D","I","D","G",
                           "E","I",
                           "F","G","F","J",
                           "G","H","G","J",
                           "H","I","H","J",
                           "I","J"), directed = TRUE)
  after_10_days.i <- rbinom(n = 22,size = 1,prob = 1-exp(-10*theta.g))
  graph_after_10_days.i <- delete.edges(graph = graph.zero_day.i,edges = which(after_10_days.i==1))
  paths_A_J_after_10_days.i <- all_simple_paths(graph = graph_after_10_days.i,from = "A","J")
  if(identical(paths_A_J_after_10_days.i,list())){
    h_x.result_vec.c[i] <- 1*prod( dbinom(x = after_10_days.i,size = 1,prob = 1-exp(-10*theta.f)) / dbinom(x = after_10_days.i,size = 1,prob = 1-exp(-10*theta.g)))
  }
  
}

```

```{r,echo=FALSE}
cat(sprintf("Expected value estimator = %s", round(mean(h_x.result_vec.c),4)),
    sprintf("Varinace estimator = %s", round(var(h_x.result_vec.c),4)),
    sep = "\n")
```

```{r}
est.int <- cumsum(h_x.result_vec.c)/(1:15000)
est.err <- sqrt( cumsum( (h_x.result_vec.c-est.int)^2 )) / (1:15000)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2, ylim = mean(h_x.result_vec.c)+100*est.err[15000]*c(-1,1))
lines(est.int + 2*est.err, col = "gold", lwd = 2)
lines(est.int - 2*est.err, col = "gold", lwd = 2)
```


## d ## 

```{r}
theta.f <- 0.005
theta.g <- 0.02

h_x.result_vec.d <- numeric(50000)

for (i in 1:50000) {
  graph.zero_day.i <- make_graph(c("A", "B", "A", "C","A", "D","A", "E",
                           "B","C","B","F",
                           "C","F","C","G","C","D","C","H",
                           "D","E","D","H","D","I","D","G",
                           "E","I",
                           "F","G","F","J",
                           "G","H","G","J",
                           "H","I","H","J",
                           "I","J"), directed = TRUE)
  after_10_days.i <- rbinom(n = 22,size = 1,prob = 1-exp(-10*theta.g))
  graph_after_10_days.i <- delete.edges(graph = graph.zero_day.i,edges = which(after_10_days.i==1))
  paths_A_J_after_10_days.i <- all_simple_paths(graph = graph_after_10_days.i,from = "A","J")
  if(identical(paths_A_J_after_10_days.i,list())){
    h_x.result_vec.d[i] <- 1*prod( dbinom(x = after_10_days.i,size = 1,prob = 1-exp(-10*theta.f)) / dbinom(x = after_10_days.i,size = 1,prob = 1-exp(-10*theta.g)))
  }
  
}

```

```{r,echo=FALSE}
cat(sprintf("Expected value estimator = %s", mean(h_x.result_vec.d)),
    sprintf("Varinace estimator = %s", var(h_x.result_vec.d)),
    sep = "\n")
```

```{r}
est.int <- cumsum(h_x.result_vec.d)/(1:50000)
est.err <- sqrt( cumsum( (h_x.result_vec.d-est.int)^2 )) / (1:50000)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2)

plot(est.int, xlab = "Mean and Error range", ylab = "", type = "l", lwd = 2, ylim = mean(h_x.result_vec.d)+100*est.err[50000]*c(-1,1))
lines(est.int + 2*est.err, col = "gold", lwd = 2)
lines(est.int - 2*est.err, col = "gold", lwd = 2)
```

```{r,echo=FALSE}
summary_table <- matrix(data =
                          c(mean(h_x.result_vec.a),var(h_x.result_vec.a),
                            mean(h_x.result_vec.b),var(h_x.result_vec.b),
                            mean(h_x.result_vec.c),var(h_x.result_vec.c),
                            mean(h_x.result_vec.d),var(h_x.result_vec.d)),
                        nrow = 2,
                        ncol = 4,
                        dimnames = list(c("Expected value estimator","Varinac value estimator"),c("A","B","C","D")))

summary_table

```

It can be seen that the difference between section A and section B is that when the probability of system failure is so low, it is more "expensive" to sample and get a "good" estimation for expectation.
When in sections C and D we used "IS" method, we actually sampled from a similar distribution but with a larger parameter. That is, we have created a situation where there is a greater chance of seeing a system failure. As a result, the samples were "cheaper". This can be seen by looking at the variance estimations: IS's variance estimations are significantly smaller than the non-IS variance estimations.



Q2.D.1
h(u) = -log(1-u)/theta
u gets values between 0 and 1. so, (1-u) gets smaller values as u increases.
Giving a small value to a log function gets a smaller value and multiplying by minus 1/theta gives us a larger value (we know theta is positive).
Therefore, our function is a monotonically increasing function in u values.

```{r}
#theta <- 0.02

#h_x.result_vec.2.d <- numeric(15000)

for (i in 1:15000) {
  graph.zero_day.i <- make_graph(c("A", "B", "A", "C","A", "D","A", "E",
                           "B","C","B","F",
                           "C","F","C","G","C","D","C","H",
                           "D","E","D","H","D","I","D","G",
                           "E","I",
                           "F","G","F","J",
                           "G","H","G","J",
                           "H","I","H","J",
                           "I","J"), directed = TRUE)
  u_vec <- runif(22)
  after_10_days.i <- -log(1-u_vec)/theta
  graph_after_10_days.i <- delete.edges(graph = graph.zero_day.i,edges = which(after_10_days.i<=10))
  paths_A_J_after_10_days.i <- all_simple_paths(graph = graph_after_10_days.i,from = "A","J")
  if(identical(paths_A_J_after_10_days.i,list())){
    h_x.result_vec.2.d[i] <- 1
  }
  
}


```
Q2.D.2

we will use antithetic sampling by using the following h functions:
- h(u) = -log(1-u)/theta   - monotonically increasing function in u values.
- h(1-u) = -log(1-(1-u))/theta = -log(u)/theta   - monotonically decreasing function in u values.

```{r}
theta <- 0.02

h_x.result_vec.2.d.inc <- numeric(15000)
h_x.result_vec.2.d.dec <- numeric(15000)

for (i in 1:15000) {
  graph.zero_day.i <- make_graph(c("A", "B", "A", "C","A", "D","A", "E",
                           "B","C","B","F",
                           "C","F","C","G","C","D","C","H",
                           "D","E","D","H","D","I","D","G",
                           "E","I",
                           "F","G","F","J",
                           "G","H","G","J",
                           "H","I","H","J",
                           "I","J"), directed = TRUE)
  u_vec <- runif(22)
  after_10_days.i.inc <- -log(1-u_vec)/theta
  after_10_days.i.dec <- -log(u_vec)/theta
  
  graph_after_10_days.i.inc <- delete.edges(graph = graph.zero_day.i,edges = which(after_10_days.i.inc<=10))
  graph_after_10_days.i.dec <- delete.edges(graph = graph.zero_day.i,edges = which(after_10_days.i.dec<=10))
  
  paths_A_J_after_10_days.i.inc <- all_simple_paths(graph = graph_after_10_days.i.inc,from = "A","J")
  paths_A_J_after_10_days.i.dec <- all_simple_paths(graph = graph_after_10_days.i.dec,from = "A","J")
  
  
  if(identical(paths_A_J_after_10_days.i.inc,list())){
    h_x.result_vec.2.d.inc[i] <- 1
  }
  if(identical(paths_A_J_after_10_days.i.dec,list())){
    h_x.result_vec.2.d.dec[i] <- 1
  }
  
}



```

var(AS_est) = var(0.5(mu1_est + mu2_est)) = (var(mu1_est)+var(mu2_est)+2*cov(mu1_est,mu2_est))/4
```{r,echo=FALSE}
cat(sprintf("Antithetic sampling expected value estimator = %s", round(0.5*(mean(h_x.result_vec.2.d.inc) + mean(h_x.result_vec.2.d.dec)),4)),
    sprintf("Antithetic sampling varinace estimator = %s", round((var(h_x.result_vec.2.d.inc) + var(h_x.result_vec.2.d.dec) + 2*cov(h_x.result_vec.2.d.inc,h_x.result_vec.2.d.dec))/4,4)),
    sep = "\n")

```
Comparison between Antithetic sampling estimator and Importance sampling:
```{r,echo=FALSE}
cat(sprintf("Antithetic sampling expected value estimator = %s", round(0.5*(mean(h_x.result_vec.2.d.inc) + mean(h_x.result_vec.2.d.dec)),4)),
    sprintf("Antithetic sampling varinace estimator = %s", round((var(h_x.result_vec.2.d.inc) + var(h_x.result_vec.2.d.dec) + 2*cov(h_x.result_vec.2.d.inc,h_x.result_vec.2.d.dec))/4,4)),
    sprintf("Importance sampling expected value estimator = %s", round(mean(h_x.result_vec.a),4)),
    sprintf("Importance sampling varinace estimator = %s", round(var(h_x.result_vec.a),4)),
    sep = "\n")


```
We can see we got similar expected value estimator, but smaller varince for the Antithetic sampling estimator.

Q2.D.3
```{r}


# antithetic sampling estimator
result_vec_d3 <- numeric(10000)

x_norm_vec <- rnorm(10000)
h_x_plus <- (exp(x_norm_vec) + 7)^(x_norm_vec/3)
h_x_minus <- (exp(-x_norm_vec) + 7)^(-x_norm_vec/3)

0.5*(mean(h_x_plus)+mean(h_x_minus))
(var(h_x_plus) + var(h_x_minus) + 2*cov(h_x_plus,h_x_minus)) / 4


# monte carlo estimator
h_x_monte_carlo <- (exp(x_norm_vec) + 7)^(x_norm_vec/3)

mean(h_x_monte_carlo)
var(h_x_monte_carlo)


```
we got that the variance of antithetic sampling estimator is 4 times smaller than the monte carlo estimator
